---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
# Random walks
## Task
Write a computer program that generates random walks (RWs) as discussed in the lecture with N steps on a discrete lattice in two dimensions. Determine the average end-to-end length R of the walks as a function of the number of steps N.

## Solution
Define a function that simulates a single random walk with *N* steps.
```{r}
walk <- function(N){
  # Simulate a single random walk with N steps
  
  x <- c(0,0) #starting position
  # Generate N directions where (0, 1, 2, 3) correspond to (left, down, right, up)
  steps <- ceiling(runif(N, min=-1, max=3))
  
  for(step in steps){
    # Choosing x/y direction is based on the number being (un)even
    # Choosing sign is based on the number being greater/smaller than two
    x <- x + (2*(step%/%2)-1) * c((step+1)%%2, step%%2)
  }
  return(x)
}
```

Average the end distance to the origin for all numbers of steps up to *N* *n_tries* times (as specified in the code).

```{r}
N <- 2e2
n_tries <- 1e2

avg <- rep(0, time=N)

for(f1 in 1:N){
  tries <- rep(0, time=n_tries)
  
  for(f2 in 1:n_tries){
    tmp <- walk(f1)
    tries[f2] <- sqrt(tmp[1]**2 + tmp[2]**2)
  }
  avg[f1] <- sum(tries)/n_tries
}

x <- seq(from=1, to=N, by=1)
sqrt <- 0.9*sqrt(x)
plot(x=x, y=avg, pch=".", cex=3,
     main="Average end-to-end length", xlab="number of steps N", ylab="average distance R(N)")
lines(x=x, y=sqrt, col="red")
```
It turns out that the end-to-end length can roughly be parameterised by $$R(N) \approx 0.9 \sqrt{N}$$ (shown in the plot as red line).